Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-12-21 ms.tex
Phil's edits from final read-through

Approximate analytic solutions for resonant scattering exist in certain limits. \citet{1973MNRAS.162...43H} showed that when most of the radiation is out on the damping wings, the transfer equation reduces to the Poisson equation. However, their solution contains an incorrect treatment of the boundary condition. To our knowledge, the errors introduced by this treatment have never been quantified. They attempt a separation of variables as $J(\tau,\sigma) = \theta(\tau) j(\sigma)$ in spatial variable $\tau$ and frequency variable $\sigma$ (their Equations 16 and 23). The solutions for the eigenfunctions $\theta(\tau)$ and $j(\sigma)$ then depend explicitly on the separation constant $\lambda$. In order to satisfy the boundary conditions, the separation constant is shown to satisfy an eigenvalue equation of the form
\be
\lambda \tan(\lambda B) & = & \frac{3}{2} \phi \Delta,
\label{eq:evalue}
\ee
where $2B$ is the slab optical depth at line center, $\phi$ is the line profile, and $\Delta$ is the Doppler width. The key point is that the line profile depends on one of the coordinates: frequency. This causes the eigenvalues of the separation constant to be frequency-dependent. Thus, the separation ``constant'' is not constant, and the function does not satisfy the Poisson equation since the frequency derivatives will act on the separation ``constant", giving extra terms. In the limit of large optical depth $B$, they approximate the eigenvalues as $\lambda_n B \simeq \pi (n-1/2)$, which gives zero mean intensity at the surface. Their Equation 34 subsequently allows $\lambda$ to have a small deviation from the above expression, which is explicitly frequency dependent. This allows a nonzero intensity at the surface, but at the cost of rendering the separation of variables assumption invalid. Our treatment, using the correct boundary condition, quantifies the errors in this ansatz.

@@

Consider a sphere of radius $R$ with uniform density $n_{\rm sc}$, luminosity $L$, and line-center optical depth $\tau_0$ containing a point source of photons. We aim to find the intensity within the sphere as a function of radius and photon frequency. The point source is assumed to be a delta function in space and photon frequency. Photons of frequency $\nu$ near the line center frequency $\nu_0$ are considered. The photon frequency of the source is $\nu_s$. The Doppler width is $\Delta = \nu_0 v_{\rm th}/c$, where $v_{\rm th}=\sqrt{2k_{\rm B}T/m_{\rm H}}$ is the thermal speed of hydrogen atoms of mass $m_{\rm H}$ and temperature $T$, and $c$ is the speed of light. The photon frequency in Doppler widths is $x = (\nu-\nu_0)/\Delta$, and $x_s = (\nu_s - \nu_0)/\Delta$ is the corresponding source frequency. For upper-state de-excitation rate $\Gamma$, the ratio of natural to Doppler broadening is $a=\Gamma/(4\pi \Delta)$. For the \lya transition and T=$10^4$ K, $a = 4.72\times 10^{-4}$. $\mathcal{H}(x,a)$ is the Voigt function, and the Voigt line profile is $\phi = \mathcal{H}(x,a)/(\sqrt{\pi} \Delta)$, which is normalized as $\int d\nu\, \phi(\nu) = 1$. The line center optical depth is $\tau_0 = kR/(\sqrt{\pi}\Delta)$, where $k = n_{\rm sc} \pi e^2 f/(m_e c)$. Here, $e$ and $m_e$ are the charge and mass of the electron, and $f$ is the oscillator strength of the transition, which is 0.4162 for \lya \citep{1986rpa..book.....R}.

@@

\be
H_{\rm d} & = & - \frac{1}{3k\phi} \frac{\partial J_d}{\partial r}
=  \left( \frac{1}{3k\phi} \right) 
\left( \frac{ \sqrt{6}k^3L }{ 8\pi^3 \Delta^4} \right)
\left( \frac{k r/\Delta}{ \left[ (kr/\Delta)^2 + (\sigma-\sigma_{\rm s})^2 \right]^2 } \right).
\label{eq:Hd}
\ee
This solution is useful as a simple analytic formula. However, it is not a good approximation to the true solution, as it is too large at $r=R$ by a factor of $J_{\rm d}(R,\sigma)/ H_{\rm d}(R,\sigma) \sim a\tau_0/x^2 \sim (a\tau_0)^{1/3} \gg 1$ and does not adhere to the correct boundary condition. This solution is included in Fig. \ref{fig:sol_mc_residual_0} for illustration.

A better approximation to the true solution has been derived by \citet{2006ApJ...649...14D}, who generalized the closed-form solution in slab geometry found in \citet{1990ApJ...350..216N}. It satisfies a $J=0$ boundary condition at $r=R$. Again, we generalize their solution to allow emission at frequency $\sigma_{\rm s}$ away from line center. The result can be written as a sum over spatial modes,
\be \label{eq:J0_sum}
J_0 = \frac{\sqrt{6}L}{16\pi \Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\frac{\sin{\kappa_n r}}{\kappa_n r}\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
and
\be \label{eq:H0_sum}
H_0 = - \frac{1}{3k\phi} \frac{\partial J_0}{\partial r} = -\frac{1}{3k\phi}\frac{\sqrt{6}L}{16\pi\Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\left(\frac{\cos{\kappa_n r}}{r} - \frac{\sin{\kappa_n r}}{\kappa_n r^2}\right)\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
where $\kappa_n=n\pi/R$. These can be summed to give the closed form expressions
\be
J_0 & = & \frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right)
\label{eq:J0}
\ee
and
\be
H_0 & = &\frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr^2}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right. \nonumber \\ & & \left. - \left( \frac{\pi r}{R} \right)
\frac{ \cos(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
+ \left( \frac{\pi r}{R} \right)
\frac{ \sin^2(\pi r/R) }{ \left[ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R) \right]^2 }
\right).
\label{eq:H0}
\ee
These solutions agree with Equations (\ref{eq:Jd}) and (\ref{eq:Hd}) when the arguments of the trigonometric and hyperbolic functions are small. Again $J_0 \gg H_0$, except near $r=R$, where it goes to zero. The flux at $r=R$ can be written
\be
\nonumber
H_0(R, \sigma) & = & - \frac{1}{3k\phi}
\frac{\sqrt{6}L}{16\pi \Delta}
\frac{1}{R^3}
\sum_{n=1}^{\infty} 
n (-1)^n \exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)}\\
& = &  \frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi \Delta}
\frac{1}{R^3}
\left( 
\frac{ 1 }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] +1 }
\right).
\label{eq:H0surf}
\ee
Equation (\ref{eq:H0surf}) will be shown to be a better approximation to the solution than Equation (\ref{eq:Hd}). It is still valid near the delta function at $r=0$, but is also a better approximation at $r=R$. $J_0$ decreases exponentially rather than as a power-law in frequency as it does for $J_d$, giving a much smaller flux on the line wings as compared to the divergent solution.

In order to enforce the boundary conditions, a different solution method is attempted here, namely a continuous Fourier expansion in the frequency variable $\sigma$. The solution of this problem is split into two pieces: $J_0$ which includes the delta function source and satisfies $J=0$ at $r=R$, and $J_{\rm bc}$ which allows the boundary condition $J=\sqrt{3}H$ to be satisfied at $r=R$. The total solution is
\be
J(r,\sigma) & = & J_0(r,\sigma) + J_{\rm bc}(r,\sigma)
\ee
and
\be \label{eq:totalflux}
H(r,\sigma) & = & H_0(r,\sigma) + H_{\rm bc}(r,\sigma).
\ee
The additional term $J_{\rm bc}$ must then be a solution of the homogeneous equation
\be \label{eq:diffeq}
\frac{\partial^2J_{\rm bc}}{\partial r^2} + \frac{2}{r} \frac{\partial J_{\rm bc}}{\partial r}
+ \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J_{\rm bc}}{\partial \sigma^2} &= & 0
\ee
with no delta function source term, and it must allow the boundary conditions to be satisfied at the surface. Since $J_0(R,
\sigma)=0$, the surface boundary condition becomes
\be
J_{\rm bc}(R,\sigma) - \sqrt{3} H_{\rm bc}(R,\sigma) & = 
\sqrt{3} H_0(R,\sigma).
\label{eq:bc2}
\ee
Inserting a frequency dependence $J_{\rm bc} \propto e^{is\sigma}$, for ``wavenumber" $s$, gives the equation for modified spherical Bessel functions of the first kind, $i_0(z)=\sinh(z)/z$ for the radial dependence. The solution can then be represented as
\be
J_{\rm bc}(r,\sigma) & = & 
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\frac{i_0(krs/\Delta)}{i_0(kRs/\Delta)},
\label{eq:Jbc}
\ee
where $A(s)$ is the Fourier amplitude. Inserting Equation (\ref{eq:Jbc}) into Equation (\ref{eq:bc2}) leads to the following equation for the Fourier amplitudes,
\be
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s)
\left[ 1 + \left( \frac{s}{\sqrt{3} \Delta \phi} \right) \left( \frac{i_0^\prime(kRs/\Delta)}{i_0(kRs/\Delta)} \right) \right]
& = & \sqrt{3} H_0(R,\sigma).
\label{eq:bc3}
\ee
Discretization of Equation (\ref{eq:bc3}) for frequency variables $\sigma_i$ and wavenumbers $s_j$
leads to a set of coupled linear equations for the $A(s_j)$. We use equally-spaced points $\delta \sigma = 2\sigma_{\rm max}/(N-1)$ and $\delta s = 2\pi/(N\delta \sigma)$, where $N$ is the number of points for each grid. The result is a linear system to solve for the $A(s)$. The maximum frequency is set as $\sigma_{\rm max} = {\rm constant} \times \tau_0$, for a large enough constant that the end of the frequency grid is at such small intensities that it does not affect the solution except close to the boundaries. The number of points was increased until the solution was well-resolved near line center, and only became inaccurate close to the boundaries. Given the Fourier amplitudes $A(s)$, $J_{\rm bc}$ is computed using Equation (\ref{eq:Jbc}), and the flux is given by
\be
H_{\rm bc}(r,\sigma) & = & -\frac{1}{3k\phi}
\frac{\partial J_{\rm bc}(r,\sigma)}{\partial r}
= -\frac{1}{3k\phi}
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\left( \frac{ks}{\Delta} \right) 
\left( \frac{i_0^\prime(krs/\Delta)}{i_0(kRs/\Delta)} \right).
\label{eq:Hbc}
\ee
The Bessel functions are finite at the center and rise steeply toward the surface when $kRs/\Delta \gg 1$.

@@

We now estimate the scaling of $H_{\rm bc}$ with $\tau_0$. In the limit $J_{\rm bc} \gg H_{\rm bc}$, we find that $J_{\rm bc} \approx \sqrt{3} H_0$ from Equation (\ref{eq:bc2}). We  calculate $H_{\rm bc}$ from $J_{\rm bc}$ using Equation (\ref{eq:Hbc}) as
\be
H_{\rm bc}(R, \sigma) \approx \frac{1}{\sqrt{3}k\phi}\frac{ks}{\Delta}H_0\ {\sim}\ H_0 s \frac{x^2}{a}\ {\sim}\ H_0 \frac{1}{\tau_0}\frac{(a\tau_0)^{2/3}}{a}\ {\sim}\ H_0 (a\tau_0)^{-1/3},
\ee
where we have used $s\ {\sim}\ 1/\sigma\ {\sim}\ 1/\tau_0$ so that
\be \label{eq:hbc_scaling}
\frac{H_{\rm bc}(R, \sigma)}{H_0(R, \sigma)} \propto (a\tau_0)^{-1/3}.
\ee
At large $\tau_0$, it is expected that the correction term will be small, but it will become increasingly important as $\tau_0$ decreases. Our solution of the transfer equation is only valid when the peaks of the spectral energy distribution lie well outside of the Doppler core, e.g., for large $\tau_0$. Thus, the value of $x$ at which the Doppler and Lorentzian components of the line profile are equal is $x_{\rm cw}=3.3$. Setting $x_{\rm cw} = x_{\rm peak}$ and solving for $\tau_0$ gives the value at which the peak of the spectrum falls at the Doppler core boundary, which is $\tau_{\rm cp} \approx 10^5$ (``core-peak'' optical depth). Hence $H_{\rm bc}/H_0 {\sim} (\tau_{\rm cp}/\tau_0)^{1/3}$ is large at $\tau_0 \leq \tau_{\rm cp}$ and decreases relatively slowly as $\tau_0$ increases. Additionally, the optical depth at $x_{\rm peak}$ is proportional to $(a\tau_0)^{1/3}$, so photons here become optically thin when $a\tau_0 {\sim} 1$.

\begin{figure}
    \centering
    \includegraphics{final_residual.pdf}
    \caption{Spectrum $P(x)$ vs. frequency $x$. The analytic solutions for the outgoing spectrum are calculated at a line center optical depth of $\tau_0 = 1 \times 10^7$ with $x_s = 0.0$. The top panel is linear scale, the middle panel is log scale with $|\rm H_{bc}|$ shown instead of $\rm H_{bc}$, and the bottom panel is the residual of each solution with Monte Carlo.} 
    \label{fig:sol_mc_residual_0}
\end{figure}

\subsection{Comparison to Monte Carlo}
The Monte Carlo method is used to solve the transfer equation numerically in order to compare the analytic approximation to an ``exact'' solution. This method is valid at all $\tau_0$, being restricted only by its computational demand which grows proportionally to the number of photons used and $\tau_0$. For each simulation, a total of ${\sim}10^6$ photon packets are initialized at a monochromatic source frequency $x_s$ and are allowed to propagate through the sphere until escaping, at which point their positions, outgoing angles, and escape frequencies are tabulated to obtain the spectrum at the surface of the spherical simulation domain. A constant temperature of $T=10^4\ \rm K$ is set for the gas. Frequency redistribution is calculated at each scattering, including the effects of recoil. In the comparisons shown in this section, the raw photon data is binned in frequency to obtain spectra. Further details of the Monte Carlo implementation are discussed in \citet{2017ApJ...851..150H}.

@@

In Figure \ref{fig:sol_mc_residual_0}, the Monte Carlo escape probability as a function of frequency is shown along with that of the solutions $H_{\rm d}$, $H_0$, and $H_0 + H_{\rm bc}$ for an optical depth of $\tau_0 = 10^7$ and photons emitted at line center $\rm x_s = 0$.  The $H_{\rm bc}$ term is negative at the peak of the spectrum and positive in the line wing such that, when added to $H_0$, it corrects for the apparent excess of flux in the peaks of the spectrum. The solution with the correct frequency-dependent boundary condition enforced, $H_0 + H_{\rm bc}$, has lower residuals to Monte Carlo results than the other solutions, especially in the line wing. The boundary term corrects the deficit of $H_0$ in the line wings, further improving agreement with the numerical result. Note that the residuals to the $H_0$ solution are a close match to the $H_{\rm bc}$ term, since the Monte Carlo represents the ``true'' solution, $H$, and $H_{\rm bc} = H - H_0$. It is evident that the divergent solution $H_{\rm d}$ fails in the line wings.

 \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{tau_threepanel.pdf}
    \caption{The same as Figure \ref{fig:sol_mc_residual_0}, but for $\tau_0 = 10^5, 10^6, 10^7$. The $x$ and $y$ axes are scaled by $(a\tau_0)^{1/3}$. Photons are initialized with $\rm x_s = 0.0$ in each case.}
    \label{fig:sol_mc_tau}
\end{figure}

@@

In Figure \ref{fig:sol_mc_tau} we show the solutions alongside Monte Carlo for $\tau_0=10^5, 10^6$, and $10^7$. From Equation (\ref{eq:hbc_scaling}), the size of the term $H_{\rm bc}$ should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. Indeed, agreement between Equation (\ref{eq:H0surf}) and the Monte Carlo points in Fig. \ref{fig:sol_mc_tau} improves as $\tau_0$ increases, with $H_{\rm bc}$ providing a fractionally smaller correction to $H_0$. One factor of $(a\tau_0)^{1/3}$ has been scaled out of the x-axis such that the peaks of the distributions are horizontally aligned. This scaling has also been applied to the y-axis to preserve normalization of the escape probability. At lower $\tau_0$, the scattering of photons within the Doppler core of the line becomes important, but our analytic solution does not include this effect. The effects of line core scattering can be seen in the Monte Carlo data for $\tau_0=10^5$ and, to a lesser extent, $
\tau_0=10^6$.

>>>>

Approximate analytic solutions for resonant scattering exist in certain limits. \citet{1973MNRAS.162...43H} showed that when most of the radiation is out on the damping wings, the transfer equation reduces to the Poisson equation. However, their solution uses an ansatz to handle the boundary condition. To our knowledge, the errors introduced by this treatment have never been quantified. They attempt a separation of variables as $J(\tau,\sigma) = \theta(\tau) j(\sigma)$ in spatial variable $\tau$ and frequency variable $\sigma$ (their Equations 16 and 23). The solutions for the eigenfunctions $\theta(\tau)$ and $j(\sigma)$ then depend explicitly on the separation constant $\lambda$. In order to satisfy the boundary conditions, the separation constant is shown to satisfy an eigenvalue equation of the form
\be
\lambda \tan(\lambda B) & = & \frac{3}{2} \phi \Delta,
\label{eq:evalue}
\ee
where $2B$ is the slab optical depth at line center, $\phi$ is the line profile, and $\Delta$ is the Doppler width. The key point is that the line profile depends on one of the coordinates: frequency. This causes the eigenvalues of the separation constant to be frequency-dependent. Thus, the separation ``constant'' is not constant, and the function does not satisfy the Poisson equation since the frequency derivatives will act on the separation ``constant", giving extra terms. In the limit of large optical depth $B$, they approximate the eigenvalues as $\lambda_n B \simeq \pi (n-1/2)$, which gives zero mean intensity at the surface. Their Equation 34 subsequently allows $\lambda$ to have a small deviation from the above expression, which is explicitly frequency dependent. This allows a nonzero intensity at the surface, but at the cost of rendering the separation of variables assumption invalid. Our treatment, using the correct boundary condition, quantifies the errors in this ansatz.

@@

Consider a sphere of radius $R$ with uniform density $n_{\rm sc}$, luminosity $L$, and line-center optical depth $\tau_0$, containing a point source of photons. We aim to find the intensity within the sphere as a function of radius and photon frequency. The point source is assumed to be a delta function in space and photon frequency. Photons of frequency $\nu$ near the line center frequency $\nu_0$ are considered. The photon frequency of the source is $\nu_s$. The Doppler width is $\Delta = \nu_0 v_{\rm th}/c$, where $v_{\rm th}=\sqrt{2k_{\rm B}T/m_{\rm H}}$ is the thermal speed of hydrogen atoms of mass $m_{\rm H}$ and temperature $T$, and $c$ is the speed of light. The photon frequency in Doppler units is $x = (\nu-\nu_0)/\Delta$, and $x_s = (\nu_s - \nu_0)/\Delta$ is the corresponding source frequency. For upper-state de-excitation rate $\Gamma$, the ratio of natural to Doppler broadening is $a=\Gamma/(4\pi \Delta)$. For the \lya transition and T=$10^4$ K, $a = 4.72\times 10^{-4}$. $\mathcal{H}(x,a)$ is the Voigt function, and the Voigt line profile is $\phi = \mathcal{H}(x,a)/(\sqrt{\pi} \Delta)$, which is normalized as $\int d\nu\, \phi(\nu) = 1$. The line center optical depth is $\tau_0 = kR/(\sqrt{\pi}\Delta)$, where $k = n_{\rm sc} \pi e^2 f/(m_e c)$. Here, $e$ and $m_e$ are the charge and mass of the electron, and $f$ is the oscillator strength of the transition, which is 0.4162 for \lya \citep{1986rpa..book.....R}.

@@

\be
H_{\rm d} & = & - \frac{1}{3k\phi} \frac{\partial J_d}{\partial r}
=  \left( \frac{1}{3k\phi} \right) 
\left( \frac{ \sqrt{6}k^3L }{ 8\pi^3 \Delta^4} \right)
\left( \frac{k r/\Delta}{ \left[ (kr/\Delta)^2 + (\sigma-\sigma_{\rm s})^2 \right]^2 } \right).
\label{eq:Hd}
\ee
This solution is useful as a simple analytic formula. However, it is not a good approximation to the true solution, as it is too large at $r=R$ by a factor of $J_{\rm d}(R,\sigma)/ H_{\rm d}(R,\sigma) \sim a\tau_0/x^2 \sim (a\tau_0)^{1/3} \gg 1$ and does not adhere to the correct boundary condition. This solution is included in Figure \ref{fig:sol_mc_residual_0} for illustration.

A better approximation to the true solution has been derived by \citet{2006ApJ...649...14D}, who generalized the closed-form solution in slab geometry found in \citet{1990ApJ...350..216N}. It satisfies a $J=0$ boundary condition at $r=R$. Again, we generalize their solution to allow emission at frequency $\sigma_{\rm s}$ away from line center. The result can be written as a sum over spatial modes,
\be \label{eq:J0_sum}
J_0 = \frac{\sqrt{6}L}{16\pi \Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\frac{\sin{\kappa_n r}}{\kappa_n r}\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
and
\be \label{eq:H0_sum}
H_0 = - \frac{1}{3k\phi} \frac{\partial J_0}{\partial r} = -\frac{1}{3k\phi}\frac{\sqrt{6}L}{16\pi\Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\left(\frac{\cos{\kappa_n r}}{r} - \frac{\sin{\kappa_n r}}{\kappa_n r^2}\right)\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
where $\kappa_n=n\pi/R$. These can be summed to give the closed form expressions
\be
J_0 & = & \frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right)
\label{eq:J0}
\ee
and
\be
H_0 & = &\frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr^2}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right. \nonumber \\ & & \left. - \left( \frac{\pi r}{R} \right)
\frac{ \cos(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
+ \left( \frac{\pi r}{R} \right)
\frac{ \sin^2(\pi r/R) }{ \left[ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R) \right]^2 }
\right).
\label{eq:H0}
\ee
These solutions agree with Equations (\ref{eq:Jd}) and (\ref{eq:Hd}) when the arguments of the trigonometric and hyperbolic functions are small. Again $J_0 \gg H_0$, except near $r=R$, where it goes to zero. The flux at $r=R$ can be written
\be
\nonumber
H_0(R, \sigma) & = & - \frac{1}{3k\phi}
\frac{\sqrt{6}L}{16\pi \Delta}
\frac{1}{R^3}
\sum_{n=1}^{\infty} 
n (-1)^n \exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)}\\
& = &  \frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi \Delta}
\frac{1}{R^3}
\left( 
\frac{ 1 }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] +1 }
\right).
\label{eq:H0surf}
\ee
Equation (\ref{eq:H0surf}) will be shown to be a better approximation to the solution than Equation (\ref{eq:Hd}). It is still valid near the delta function at $r=0$, but is also a better approximation at $r=R$. $J_0$ decreases exponentially, rather than as a power-law in frequency as it does for $J_d$, giving a much smaller flux on the line wings as compared to the divergent solution.

In order to enforce the boundary conditions, a different solution method is attempted here, namely a continuous Fourier expansion in the frequency variable $\sigma$. The solution of this problem is split into two pieces: $J_0$ which includes the delta function source and satisfies $J=0$ at $r=R$, and $J_{\rm bc}$ which allows the boundary condition $J=\sqrt{3}H$ to be satisfied at $r=R$. The total solution is
\be
J(r,\sigma) & = & J_0(r,\sigma) + J_{\rm bc}(r,\sigma)
\ee
and
\be \label{eq:totalflux}
H(r,\sigma) & = & H_0(r,\sigma) + H_{\rm bc}(r,\sigma).
\ee
The additional term $J_{\rm bc}$ must then be a solution of the homogeneous equation
\be \label{eq:diffeq}
\frac{\partial^2J_{\rm bc}}{\partial r^2} + \frac{2}{r} \frac{\partial J_{\rm bc}}{\partial r}
+ \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J_{\rm bc}}{\partial \sigma^2} &= & 0
\ee
with no delta function source term, and it must allow the boundary conditions to be satisfied at the surface. Since $J_0(R,
\sigma)=0$, the surface boundary condition becomes
\be
J_{\rm bc}(R,\sigma) - \sqrt{3} H_{\rm bc}(R,\sigma) & = 
\sqrt{3} H_0(R,\sigma).
\label{eq:bc2}
\ee
Inserting a frequency dependence $J_{\rm bc} \propto e^{is\sigma}$, for ``wavenumber" $s$, gives the equation for modified spherical Bessel functions of the first kind, $i_0(z)=\sinh(z)/z$ for the radial dependence. The solution can then be represented as
\be
J_{\rm bc}(r,\sigma) & = & 
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\frac{i_0(krs/\Delta)}{i_0(kRs/\Delta)},
\label{eq:Jbc}
\ee
where $A(s)$ is the Fourier amplitude. Inserting Equation (\ref{eq:Jbc}) into Equation (\ref{eq:bc2}) leads to the following equation for the Fourier amplitudes,
\be
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s)
\left[ 1 + \left( \frac{s}{\sqrt{3} \Delta \phi} \right) \left( \frac{i_0^\prime(kRs/\Delta)}{i_0(kRs/\Delta)} \right) \right]
& = & \sqrt{3} H_0(R,\sigma).
\label{eq:bc3}
\ee
Discretization of Equation (\ref{eq:bc3}) for frequency variables $\sigma_i$ and wavenumbers $s_j$
leads to a set of coupled linear equations for the $A(s_j)$. We use equally-spaced points $\delta \sigma = 2\sigma_{\rm max}/(N-1)$ and $\delta s = 2\pi/(N\delta \sigma)$, where $N$ is the number of points for each grid. The maximum frequency is set as $\sigma_{\rm max} = {\rm constant} \times \tau_0$, for a large enough constant that the end of the frequency grid is at such small intensities that it does not affect the solution except close to the boundaries. The number of points was increased until the solution was well-resolved near line center, and only became inaccurate close to the boundaries. Given the Fourier amplitudes $A(s)$, $J_{\rm bc}$ is computed using Equation (\ref{eq:Jbc}), and the flux is given by
\be
H_{\rm bc}(r,\sigma) & = & -\frac{1}{3k\phi}
\frac{\partial J_{\rm bc}(r,\sigma)}{\partial r}
= -\frac{1}{3k\phi}
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\left( \frac{ks}{\Delta} \right) 
\left( \frac{i_0^\prime(krs/\Delta)}{i_0(kRs/\Delta)} \right).
\label{eq:Hbc}
\ee
The Bessel functions are finite at the center and rise steeply toward the surface when $kRs/\Delta \gg 1$.

@@

We now estimate the scaling of $H_{\rm bc}$ with $\tau_0$. In the limit $J_{\rm bc} \gg H_{\rm bc}$, we find that $J_{\rm bc} \approx \sqrt{3} H_0$ from Equation (\ref{eq:bc2}). We  estimate $H_{\rm bc}$ from $J_{\rm bc}$ using Equation (\ref{eq:Hbc}) as
\be
H_{\rm bc}(R, \sigma) \approx \frac{1}{\sqrt{3}k\phi}\frac{ks}{\Delta}H_0\ {\sim}\ H_0 s \frac{x^2}{a}\ {\sim}\ H_0 \frac{1}{\tau_0}\frac{(a\tau_0)^{2/3}}{a}\ {\sim}\ H_0 (a\tau_0)^{-1/3},
\ee
where we have used $s\ {\sim}\ 1/\sigma\ {\sim}\ 1/\tau_0$ so that
\be \label{eq:hbc_scaling}
\frac{H_{\rm bc}(R, \sigma)}{H_0(R, \sigma)} \propto (a\tau_0)^{-1/3}.
\ee
At large $\tau_0$, it is expected that the correction term will be small, but it will become increasingly important as $\tau_0$ decreases. Our solution of the transfer equation is only valid when the peaks of the spectral energy distribution lie well outside of the Doppler core, e.g., for large $\tau_0$. The value of $x$ at which the Doppler and Lorentzian components of the line profile are equal is $x_{\rm cw}=3.3$. Setting $x_{\rm cw} = x_{\rm peak}$ and solving for $\tau_0$ gives the value at which the peak of the spectrum falls at the Doppler core boundary, which is $\tau_{\rm cp} \approx 10^5$ (``core-peak'' optical depth). Hence $H_{\rm bc}/H_0 {\sim} (\tau_{\rm cp}/\tau_0)^{1/3}$ is large at $\tau_0 \leq \tau_{\rm cp}$ and decreases relatively slowly as $\tau_0$ increases. Additionally, the optical depth at $x_{\rm peak}$ is proportional to $(a\tau_0)^{1/3}$, so photons here become optically thin when $a\tau_0 {\sim} 1$.

\begin{figure}
    \centering
    \includegraphics{final_residual.pdf}
    \caption{Spectrum $P(x)$ vs. frequency $x$ for $\tau_0 = 1 \times 10^7$ with $x_s = 0.0$. The legend to the right describes each line style. The top panel is linear scale, the middle panel is log scale with $|\rm H_{bc}|$ shown instead of $\rm H_{bc}$, and the bottom panel is the residual of each solution with Monte Carlo.} 
    \label{fig:sol_mc_residual_0}
\end{figure}

\subsection{Comparison to Monte Carlo}
The Monte Carlo method is used to solve the transfer equation numerically in order to compare the analytic approximation to an ``exact'' solution. This method is valid at all $\tau_0$, being restricted only by the computational demand, which grows proportionally to the number of photons used and $\tau_0$. For each simulation, a total of ${\sim}10^6$ photon packets are initialized at a monochromatic source frequency $x_s$ and are allowed to propagate through the sphere until escaping, at which point their positions, outgoing angles, and escape frequencies are tabulated to obtain the spectrum at the surface of the spherical simulation domain. A constant temperature of $T=10^4\ \rm K$ is set for the gas. Frequency redistribution is calculated at each scattering. In the comparisons shown in this section, the raw photon data is binned in frequency to obtain spectra. Further details of the Monte Carlo implementation are discussed in \citet{2017ApJ...851..150H}.

@@

In Figure \ref{fig:sol_mc_residual_0}, the Monte Carlo spectrum is shown along with that of the solutions $H_{\rm d}$, $H_0$, and $H_0 + H_{\rm bc}$ for an optical depth of $\tau_0 = 10^7$ and photons emitted at line center $\rm x_s = 0$.  The $H_{\rm bc}$ term is negative at the peak of the spectrum and positive in the line wing such that, when added to $H_0$, it corrects for the apparent excess of flux in the peaks of the spectrum. The solution with the correct frequency-dependent boundary condition enforced, $H_0 + H_{\rm bc}$, has lower residuals to Monte Carlo results than the other solutions, especially in the line wing. The boundary term corrects the deficit of $H_0$ in the line wings, further improving agreement with the numerical result. Note that the residuals to the $H_0$ solution are a close match to the $H_{\rm bc}$ term, since the Monte Carlo represents the ``true'' solution, $H$, and $H_{\rm bc} = H - H_0$. It is evident that the divergent solution $H_{\rm d}$ fails in the line wings.

 \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{tau_threepanel.pdf}
    \caption{The same as Figure \ref{fig:sol_mc_residual_0}, but for $\tau_0 = 10^5$ (top panel), $10^6$ (middle panel), and $10^7$ (bottom panel). The $x$ and $y$ axes are scaled by $(a\tau_0)^{1/3}$.}
    \label{fig:sol_mc_tau}
\end{figure}

@@

In Figure \ref{fig:sol_mc_tau} we show the solutions alongside Monte Carlo, now for three different optical depths $\tau_0=10^5, 10^6$, and $10^7$. From Equation (\ref{eq:hbc_scaling}), the size of the term $H_{\rm bc}$ should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. Indeed, agreement between Equation (\ref{eq:H0surf}) and the Monte Carlo points in Figure \ref{fig:sol_mc_tau} improves as $\tau_0$ increases, with $H_{\rm bc}$ providing a fractionally smaller correction to $H_0$. One factor of $(a\tau_0)^{1/3}$ has been scaled out of the x-axis such that the peaks of the distributions are horizontally aligned. This scaling has also been applied to the y-axis to preserve normalization of the escape probability. At lower $\tau_0$, the scattering of photons within the Doppler core of the line becomes important, but our analytic solution does not include this effect. The effects of line core scattering can be seen in the Monte Carlo data for $\tau_0=10^5$ and, to a lesser extent, $
\tau_0=10^6$.
